How has the Brazilian Amazon been constructed as a problem? An analysis of

### 2022-04-11 Draft_April_2022/Full_draft_20220401.rmd

# 5 Conclusion

# 6 References

>>>>

# 5 Discussion

# 6 Conclusion

# References

# Appendix

## Extended Methodology

This appendix details the methodology employed by the authors in the article.

## The Dataset

We update a dataset provided by Cezar (2020), which builds upon the archives of the Brazilian Presidential Library. The dataset contains all official remarks by Brazilian Presidents from 1985 to 2019. We update the dataset by scraping and adding all official speeches from 2020 and 2021. The final dataset encompasses 6130 speeches. Overall, it covers a total of 8 presidents (Sarney, Collor, Franco, Cardoso, Silva, Rousseff, Temer, and Bolsonaro) across 8 mandates (Collor and Dilma were impeached and vice presidents Franco and Temer took over).

We start by cleaning the dataset by removing all accents, non-standard Latin characters, and extra spaces. We then identify all speeches about the Amazon as a region, peoples, or forest out of the 6130 speeches. Our dictionary is composed of a single lexicon stem: “amazon”. In Portuguese, it captures terms such as “Amazonia”, “Amazonica”, “Amazonidas”, “Amazonense(s)”, “Amazonas”, among others. We chose to do so knowing that a few speeches about the Amazon might not contain the lexicon “Amazon”, for example, when the president says “the forest” or uses a pronoun, he or she could be referring to the Amazon. Also, it is likely that whenever a president mentions “deforestation” or derived vocabulary, he or she is speaking about the Amazon. However, as we cannot be sure all the time in those instances (i.e. mentions of “the forest” or “deforestation” could be referring to the Atlantic Forest or the Cerrado). We opt for the safer side and make sure speeches we gather are indeed about the Amazon as a region, forest, or people by only capturing “amazon” related lexicon. We find that 946 speeches are, at least partially, about the Amazon from the 6130.

Using the poldis R package, we extract two sentences before and two sentences after the sentence in which the stem “amazon” appears. By doing so we create our unit of analysis: an Amazonian statement. We use Amazon statements as our unit of analysis for three main reasons. First, presidential speeches can be long and cover a wide variety of topics. For example, the average word count for speeches that mention the stem “amazon” are 2355 words. Working at the level of statements allows us to identify only passages that are meaningful for our specific purpose. A single speech can have more than one Amazonian statement and this statement might be embedded within a different part of the speech. Second, it increases the number of our observations and its variety meaningfully, allowing for more specificity in our analysis. Finally, other authors adopting similar strategies, opt for picking an arbitrary number of words before and after the mention (Kentikellenis and Voeten, 2021). We opt for picking two sentences before and two sentences after, rather than words because a sentence, because sentences usually contain a cohesive idea. Opting for a word threshold, might cut a sentence in the middle and yield an incomplete idea. Nevertheless, this process yields 2014 unique Amazonian statements across the 946 speeches about the Amazon identified previously. Notice that an Amazonian statement can contain two or more matches of the stem “amazon” (i.e. two consecutive sentences in which one of the words covered by the stem appears). In this case, we get two sentences before the first match and two sentences after the last match. On average, an Amazonian statement contains 123 words.

## Operationalization

We proceed to operationalize how the Amazon is constructed as a problem. To do so, we develop and apply a qualitative codebook to Amazonian statements. We do so for two reasons. First, and foremost, we believe that a dictionary and other “bag of words” methods might not capture nuances in the texts. These nuances include Amazonian statements that negate certain aspects about the Amazon or statements in which the matched words might not be not directed at the Amazonian region, peoples, or forest. Second, we chose to look at the statements and inductively develop a codebook to verify if the insights from the literature hold within presidential speeches and, if not, add or remove codes, as well as discuss which statements should, and not, be included in each code and why.

After revising the literature (section 2), we went through a random sample of the speeches to develop the codebook. The three policy-cycles (sovereignty, economic integration, and conservation) were clearly identifiable in the speeches. However, once the coding began to take place, however, we noticed two things. One, that some of the Amazon statements were not about Amazon per se. That is, our strategy to identify Amazonian statements yielded a few false positives. These are statements that mention the stem “amazon”, but are not about the Brazilian Amazon region, people, or forest. For example, the president might greet the Governor of the Amazonas in a speech or make a reference to the Venezuelan Amazon. We coded these statements as false positives. Two, we noticed speeches often refer to social development using a different vocabulary of policies and objectives. Those references often differ from references to conservation or economic integration. Hence, we created the code category ‘social development’ as explained below. We consider this to be a finding, but rather than emerging from our analysis, it emerged from codebook development. contain a full description of each code category and the coding practices that inform our procedure.

Table 1 – Qualitative Codebook

Sovereignty
Description: This code constructs the Amazon region and/or forest as an issue of national sovereignty. We understand claims of sovereignty as a particular narrative that touches on imaginaries of external threats to territory. Relatedly, we also understand sovereignty as a particular narrative that raises concerns about wrong perspectives and criticism from foreign and non-state actors about governments’ actions related to the Brazilian Amazon. In all, it advances the view that the Amazon is Brazilian and foreign presence in the region needs to be monitored closely.

Coding practice: we apply this code to references to sovereignty that meaningfully reflect the description above. Simple mentions to the Brazilian Armed Forces, frontiers, or military technology (e.g. SIPAM-SIVAM), are not automatically, by themselves, understood as constructing the Amazon as a problem of sovereignty. When presidents question the authority or interests of someone’s arguments about Amazon, for example, we apply this code as it raises concerns about wrong perspective and criticism. When presidents claim external actors want to internationalize the Amazon territory, we apply this code as it touches upon external threats to territory.

Example: “Senhores deputados. Elo fundamental para que o Brasil realmente rume em direção à prosperidade. Queria primeiro dizer, senhor Hu Chunhua,  eu quero agradecer as palavras do seu embaixador no Brasil reconhecendo a nossa soberania sobre a região Amazônica, no episódio ocorrido há pouco, por ocasião do encontro do G7. Muito obrigado ao governo chinês. Para nós, não têm preço esse reconhecimento público e suas palavras sobre essa região tão importante para o mundo e para o Brasil.” (Bolsonaro 25/10/2019)
Economic Integration
Description: This code constructs the Amazon region and/or forest as an issue of economic integration. It advances the view that the Amazon needs to be developed and connected to the national economy. This includes expanding the agricultural frontier through incentives, creating a diverse set of infrastructure (roads, dams, internet, radio, energy), fostering differing industries (tourism, mining, cattle, agriculture and so on) through tax-free zones, as well as facilitating the exploitation of natural resources for developmental purposes.

Coding practice: We apply this code to references to economic integration that meaningfully reflect the description above. Simple mentions of development, integration, investment, and technology are not automatically coded as economic integration. However, direct mentions of the words “economic development” are coded here. We avoid overlap between this code and social development by considering integration references as targeting infrastructure rather than peoples’ rights. For example, developing infrastructure for internet, radio, and energy are coded as economic integration. However, if the presidents speak about people’s access and rights to information (internet and radio) and electricity, we code as social development. Relatedly, improving macro-structural conditions of the job market is coded as economic integration,, while people’s right to a dignified job is coded as social development. Importantly, we make the analytical choice to include micro-finance instruments as credits and loans in economic development, while grants or donations fall within social development.

Example: "Se me permitem, na Amazônia -que durante tanto tempo ficou adormecida, por falta de uma ação integrada- nós temos já algumas ações muito estruturadoras. Nós, na Amazônia, estamos fazendo uma ligação que passa por Manaus, passa por Boa Vista, Caracaraí, até chegar lá em cima, que é aquela lista vermelha, que vai lá para cima, na direção da Venezuela,é a estrada 174. Essa estrada dará possibilidade da produção da Zona Franca de Manaus ser competitiva, não aqui dentro, mas lá fora mesmo, como é a vocação natural da Zona Franca, exporte, e poderá exportar pelo Caribe.” (FHC 2/7/1997)
Social Development
Description: This code constructs the Amazon region and/or forest as an issue of social development. It advances the view that Amazon is full of citizens who should have their rights guaranteed. This refers to the construction of schools and universities (right to education), of hospitals (right to health), and of housing (right to house). This also includes guarantees of a dignified life with decent employment, access to water and sanitation, as well as access to electricity, internet, radio, and light. Finally, this includes referrals to culture and the right to vote.

Coding practice: We apply this code to references to social development that meaningfully capture the description above. Simple mentions to development are not automatically coded as social development. However, direct mentions supporting “social development” are coded here. An illustrative example relates to speeches outlining the role of the armed forces in providing health services for faraway populations, which we code as social development but not sovereignty.

Example: “Mas como o Estado não tem que ter lucro, o Estado precisa garantir a cidadania, nós achamos que o cidadão que mora às margens do rio Amazonas, a 600 Km de Manaus, ele tem que ter direito a ter luz na sua casa, a ter geladeira, a ter televisão e a ver sua novela. Já investimos 14 bilhões de reais nesse programa, em três anos e meio. Sabe quantos postes nós já colocamos? Um milhão de quilômetros de fio.” (Lula 20/11/2009)
Conservation
Description: This code constructs the Amazon region and/or forest as an issue of conservation. This narrative focuses on the value of a standing forest and of the preserved ecosystem in the region. The conservationist narrative advances the view that Amazon should be preserved, deforestation should be halted, and the practices of indigenous peoples should be maintained and fostered. It advances the view that the emission of greenhouse gasses should be halted, that renewable energy should be supported, and that protected areas should be created.

Coding practice: We apply this code to references to conservation that meaningfully capture the description above. Simple mentions of biodiesel, renewable energies, climate change, or sustainable development are not automatically coded as constructing the Amazon as a conservation problem. However, calls for more demarcation of territory are coded as constructing the Amazon as a conservation problem. While general references to tourism are coded as economic integration, references to eco-tourism are coded here.

Example: “Adotei medidas de caráter emergencial, suspendi a exportação de madeiras em toras, suspendi incentivos fiscais e creditícios na Amazônia a projetos que podiam causar danos ambientais e instituí a obrigatoriedade do licenciamento para a atividade de extração de ouro com a proibição da utilização do mercúrio. Dando início à reestruturação do sistema governamental de controle e preservação do meio ambiente, criei o Instituto Brasileiro do Meio Ambiente e Recursos Naturais, IBAMA, que está entregue à capacidade do Dr. Fernando César Mesquita.” (Sarney, 20/7/1989)
False positive
Description: statements that were assigned as Amazonian but are about the Amazon.

Example: “cumprimentar os presidentes de bancos publicos do banco do brasil alexandre abreu do bndes luciano coutinho da caixa miriam belchior do banco do nordeste marcos costa holanda do banco da amazonia valmir pedro rossi o vicepresidente do banco do brasil cesar borges da diretoria de infraestrutura queria cumprimentar as senhoras e os senhores dirigentes de agencias reguladoras dirigir um cumprimento especial ao senhor robson andrade presidente da confederacao nacional da industria por intermedio de quem cumprimento todos os empresarios das federacoes aqui presentes” (Rousseff, 9/7/2015)

Each code is mutually exclusive in its conceptualization, meaning that they cover different forms of constructing the Amazon as a problem. Each Amazonian statement, though, might be assigned to one or more codes. A statement can, for example, construct the Amazon as a problem of sovereignty and a problem of economic integration or a problem of social development and conservation, and so on. Amazonian statements, thus, can be either coded as a single code (ideal types) or one or more codes (mixed types).

With this codebook in hands, then, each one of us manually coded the same set of 1007 randomly selected Amazonian statements. This amount refers to 50% of all the Amazon Statements identified. We coded all statements separately from each other, following inter-coder reliability procedures. After reading a statement, we assigned the value of one or zero to each one of the four code categories in table 1. Inter-coder agreement for each of the four main categories was 85%, on average. Most of the disagreements were on economic integration versus social development and related to the difference in emphasizing infrastructure versus emphasizing peoples’ rights. For each non-matching code, the co-authors discussed and sorted their disagreements to finalize the putting together a training set so that a supervised machine-learning model classifies the rest.

## Supervised Machine-Learning

We use this hand-coded set to train a model that codes the remaining Amazonian statements in our text corpus using R. To select the best approach, we use the hand-coded data to compare how several supervised-machine learning methods perform in identifying false positive observations from the training set. To do so, we do some simple text pre-processing in R such as removing stopwords, punctuations, signs, and empty spaces using the TM package. Then, we split the hand coded sample into a randomly selected training set and a validation set. The training set contains 80% of the observations in the hand-coded sample (806 observations) while the validation set contains the other 20% of the observations (201 observations).

We use this training set to train models support-vector machine( SVM) models to label text data on the validation set. Table 2, below, describes how the model performs in labeling observations in the validation set in regard to accuracy and AUC scores (Area Under the Curve)..  The SVM training algorithm is a non-probabilistic binary linear classifier that classifies documents by assigning points in mapped space to maximize the width of the gap between categories (Meyer et al. 2021). SVM is a particularly appropriate machine-learning algorithm for binary text labeling since it relies on maximizing distances between two categories (see also Noble 2006). We rely on the e1071 package to run the SVM model due to its flexibility in regard to parameter choices and because it helps users tune parameters to better fit the label patterns in the text data. Once the model parameters are tuned to better fit the patterns for each code, we expect the model to perform considerably better in labeling the remaining 1007 statements automatically.

Table 2 - SVM Model Performance

Accuracy
AUC
False Positives
0.946
0.69
Sovereignty
0.866
0.59
Economic Integration
0.678
0.68
Social Development
0.792
0.66
Conservation
0.802
0.78

After the SVM models are tuned, the probability that a certain Amazonian statement belongs to a code category is returned. Statements that have a probability of 0.45 or more of belonging to a code category are labeled 1 (i.e. belonging to a category) and statements that have a probability of 0.45 or less of belonging to a code category are labeled 0 (i.e. not belonging to a category). We find that a 0.5 threshold can be too strict while a 0.4 may generate a few false positives and, for this reason, we set the threshold at 0.45. The final trained models excel to label observations in some categories in comparison to others. We also take a random sample of 101 (10% of the automatically labeled Amazonian statements to check how the model performed. We find that the model performs much better than the metrics above appear to indicate in comparison to how the authors would hand code these observations. The final models automatically label, out of the remaining 1007 statements, 84 statements as sovereignty, 437 as economic integration, 186 as social development, and 300 as conservation. The model also identified 41 false positive Amazon statements out of the 1007 labeled. Moreover, the proportions of the automatically labeled Amazonian statements for all categories in regard to the number of observations in the set, including observations that were not labeled but were not false positives (i.e. others), are similar to the proportions of the coded statements for the hand coded training set.

After merging the data altogether for analysis, we chose to delete the false positive Amazonian statements for analysis since these are not meaningful statements about the Amazon as a forest, peoples, or region. In total there were 119 false positives out of the 2014 Amazonian statements. The final dataset for analysis contains 1895 Amazonian statements. Out of these, 237 statements are labeled as national sovereignty, 810 as economic integration, 418 as social development, and 567 as environmental conservation. We then added metadata about each of the speeches from which these Amazonian statements were taken from. The metadata includes variables such as the president (speaker), the year, the speech title, and the full text for the broader speech. Lastly, we use poldis once again to extract locations from texts. These locations can represent the Brazilian state in which certain speech was given, or international country or forum (e.g. United Nations).

## Multinomial Modeling

To analyze our data, we first present a series of different plots on proportions of Amazonian statements and problem-constructions over time and by presidents. We control proportions for factors that might affect the incidence of the Amazon in speeches and portray both real and controlled curves. To test whether different problem constructions change according to location, we run a multinomial regression model in which different problem constructions (as categories) are the dependent variable and location is the independent variable. We code 4 different location categories for where speeches take place, they are: Amazonian states, non-Amazonian states, Brasilia, and International. In the model, we also control for annual deforestation rates, annual inflation rates, and election years in the model as these are variables which could mitigate the correlations between problem construction and location. Annual deforestation rates data were taken from the Satellite Monitoring of Deforestation Program (PRODES). However, since annual deforestation rates are calculated from 1988, we input the rates of 1988 to 1985, 1986, and 1987 because these rates are consistent with most estimates (see Prates and Bracha 2010). The average annual inflation data is taken from the Worldwide Inflation Data. Election years are input as a dummy variable in which election years are marked 1 and non-election years are marked 0. After the model is run, post estimation tests such as the Chi-square test shows us that the model selected is highly statistically significant in explaining variance in data in comparison to a null model (i.e. our model explains a considerable amount of the variance). The likelihood ratio test also shows that the model does not suffer with issues of multicollinearity.

## Limitations

These procedures come with its limitations. Our codebook is developed using specific Amazon related vocabulary. For example, a statement will be coded as economic integration if it is a meaningful support to the Zona Franca of Manaus or a Dam in the Amazon. However, the economy is generally a topic that presidents speak about. Hence, high incidence of economic integration in Amazonian statements can also be related to the higher importance of this problem-construction in Brazil in time. Moreover, we classify statements as Amazonian based on a dictionary composed of a single lexicon stem: “amazon”. We chose to do so knowing that a few speeches about the Amazon might not contain the lexicon “Amazon”, for example, when the president says, “the forest” or “deforestation”. Hence, we might be missing statements about Amazon that do not refer to it. However, we consider this safer as we cannot be sure that mentions of the forest or deforestation do not correspond to other biomes such as the Cerrado or the Mata Atlantica. Nevertheless, our dataset covers only what is considered an official remark. Presidents, though, give interviews, appear in debates, talk at campaign rallies, and more recently start to appear in social media. Problem-construction within presidential discourse, thus, also happens in different sites for which we do not account for in this paper.
